---
output: md_document
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)


```
<!-- badges: start -->
[![CRAN Version](https://img.shields.io/cran/v/bama?style=flat-square&color=blue&label=CRAN)](https://cran.r-project.org/package=bama)
[![GitHub Release](https://img.shields.io/github/v/release/umich-cphds/bama?include_prereleases&label=Github&style=flat-square)](https://github.com/umich-cphds/bama)
[![Travis CI](https://img.shields.io/travis/umich-cphds/bama?style=flat-square)](https://travis-ci.org/umich-cphds/bama)

# Bayesian Mediation Analysis

Perform mediation analysis in the presence of high-dimensional mediators based
on the potential outcome framework. Bayesian Mediation Analysis (BAMA),
developed by Song et al (2019) <doi:10.1111/biom.13189>, relies on two Bayesian
sparse linear mixed models to simultaneously analyze a relatively large number
of mediators for a continuous exposure and outcome assuming a small number of
mediators are truly active. This sparsity assumption also allows the extension
of univariate mediator analysis by casting the identification of active
mediators as a variable selection problem and applying Bayesian methods with
continuous shrinkage priors on the effects.

## Installation
You can install `bama` via CRAN

```{r, eval = F}
install.packages("bama")
```
Or devtools
```{r, eval = F}
devtools::install_github("umich-cphds/bama", build_opts = c())
```

The Github version may contain new features or bug fixes not yet present on
CRAN, so if you are experiencing issues, you may want to try the Github
version of the package.

If you wish to install the package via devtools, you will need a C++ compiler
installed. This can be accomplished by installing Rtools on Windows and Xcode
on MacOS.

## Example
This example is taken from the `bama` help file to help you get started using
the method. Please check the documentation of the function by typing
`?bama::bama`, and the vignette by typing `vingette("bama")` in R.

`bama` includes an example dataset, `bama.data`. It is a `data.frame` with a
numeric response `y`, numeric exposure `a` and 100 numeric mediators named
`m1, m2, ..., m100`.

We recommend using much larger numbers for `burnin` and `ndraws`, for
example (30000, 1000).

```{r}
library(bama)

Y <- bama.data$y
A <- bama.data$a

# grab the mediators from the example data.frame
M <- as.matrix(bama.data[, paste0("m", 1:100)], nrow(bama.data))

# We just include the intercept term in this example.
C1 <- matrix(1, nrow(M), 1)
C2 <- matrix(1, nrow(M), 1)
beta.m  <- rep(0, 100)
alpha.a <- rep(0, 100)

set.seed(1245)
bama.out <- bama(Y, A, M, beta.m, alpha.a, C1 = C2, C2 = C2, burnin = 1000,
                 ndraws = 100)

# Rank mediators and see summary information
head(summary(bama.out, rank = T))
```

# Reference
Song, Y, Zhou, X, Zhang, M, et al. Bayesian shrinkage estimation of high
dimensional causal mediation effects in omics studies. Biometrics. 2019;
1-11. [https://doi.org/10.1111/biom.13189](https://doi.org/10.1111/biom.13189)
